### 2024-04-29 article.tex
appropriate references added

	}
\section{Proposed Methodlogy}
\subsection{Extraction of Frames and Audio from Video}
The user can upload video files of different durations. To create proper input for our audio and vision model, first the audio is extracted from the video which is then broken down into 5 sec chunks. Moreover, the image frames are extracted from the video at the rate of 1fps. Both audio and video are seperately sent to the audio and vision model respectively which then returns the output which is displayed to the user.
The following steps are involved in temporal segmentation of audio after extraction:
\begin{itemize}
	\item {The duration of the audio is calculated.}
	      \item{The number of 5 second audio segments that can be created is calculated. }
	\item { The audio is then split into 5 second segments. If any of the segments is less than 5 seconds then they are discarded. }
\end{itemize}
The following steps are involved in the splitting of video into image frames:
\begin{itemize}
	\item {The duration of video is calculated.}
	      \item{At the rate of 1fps, image frames are extracted from the video using ffmpeg.}
\end{itemize}
\subsection{Mel-frequency cepstrum Coefficients based on DCT-II for Audio Feature Extraction
}
Mel-frequency cepstral coefficients (MFCCs) are a set of features which represents the characteristics of sound. These coefficients are generated in two steps. First, a mel-scaled spectogram is created from the audio which are then used to generate MFCCs.
Mel Spectogram is a graphic representation of a sound wave. It visualises frequency over time. The process of generating a Mel Spectogram can be explained in the following steps:
\begin{itemize}
	\item {The audio signal is broken down into short frames.}
	\item {The time signal is converted into frequency domain using a STFT.}
	      \item{The frequencies are then converted into Mel scale.}
\end{itemize}
The mel scale is a perceptual scale of pitches judged by listeners to be equal in distance from one another. The following formulae is used to convert frequencies to mel scale.
\[
	Mel(f)=1125ln(1+ \frac{f}{100})
\]
The values computed above are highly correlated, which is not ideal for machine learning tasks. To solve this, Discrete Cosine Transform-II (DCT-II) is applied to decorrelate the coefficients. The then generated coefficients are also normalized orthogonally. The choice of DCT II is primarily because of its ability to compactly represent signal energy. The number of MFCCs to return is chosen to be 256 because it was easier to later transform to appropriate size for EfficientNetB0.
\subsection{Expansion and Interpolation of Audio Data}
After MFCC transformation, the tensors are transformed again to proper dimension as required by EfficientNetB0 architecture. Following steps are performed:
\begin{itemize}
	\item The input tensor y is replicated along one dimension to create three copies, preparing it for subsequent operations.
	\item The expanded tensor is resized to a fixed size of (224, 224) using bilinear interpolation, ensuring consistent dimensions for neural network input.
	\item Any extra dimensions introduced during resizing are removed, returning the tensor to its original shape.
	\item The processed tensor is returned for further use, completing the preprocessing steps commonly applied to image data for deep learning tasks.
\end{itemize}
% \begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
% def expand_tensor_for_efficient_net(y) -> torch.Tensor:
%     input_tensor_expanded = y.expand(3, -1, -1)
%     resized_tensor = F.interpolate(input_tensor_expanded.unsqueeze(
%         0), size=(224, 224), mode='bilinear', align_corners=False)

>>>>

	}
\section{Proposed Methodlogy}
\subsection{Extraction of Frames and Audio from Video}
The user can upload video files of different durations. To create proper input for our audio and vision model, first the audio is extracted from the video which is then broken down into 5 sec chunks. Moreover, the image frames are extracted from the video at the rate of 1fps. Both audio and video are seperately sent to the audio and vision model respectively which then returns the output which is displayed to the user.
The following steps are involved in temporal segmentation of audio after extraction:
\begin{itemize}
	\item {The duration of the audio is calculated.}
	      \item{The number of 5 second audio segments that can be created is calculated. }
	\item { The audio is then split into 5 second segments. If any of the segments is less than 5 seconds then they are discarded. }
\end{itemize}
The following steps are involved in the splitting of video into image frames:
\begin{itemize}
	\item {The duration of video is calculated.}
	      \item{At the rate of 1fps, image frames are extracted from the video using ffmpeg.}
\end{itemize}
\subsection{Mel-frequency cepstrum Coefficients based on DCT-II for Audio Feature Extraction
}
Mel-frequency cepstral coefficients (MFCCs) are a set of features which represents the characteristics of sound.\cite{r9} These coefficients are generated in two steps. First, a mel-scaled spectogram is created from the audio which are then used to generate MFCCs.
Mel Spectogram is a graphic representation of a sound wave. It visualises frequency over time. The process of generating a Mel Spectogram can be explained in the following steps:
\begin{itemize}
	\item {The audio signal is broken down into short frames.}
	\item {The time signal is converted into frequency domain using a STFT.}
	      \item{The frequencies are then converted into Mel scale.}
\end{itemize}
The mel scale is a perceptual scale of pitches judged by listeners to be equal in distance from one another. The following formulae is used to convert frequencies to mel scale.
\[
	Mel(f)=1125ln(1+ \frac{f}{100})
\]
The values computed above are highly correlated, which is not ideal for machine learning tasks. To solve this, Discrete Cosine Transform-II (DCT-II) is applied to decorrelate the coefficients. The then generated coefficients are also normalized orthogonally. The choice of DCT II is primarily because of its ability to compactly represent signal energy. The number of MFCCs to return is chosen to be 256 because it was easier to later transform to appropriate size for EfficientNetB0.
\subsection{Expansion and Interpolation of Audio Data}
After MFCC transformation, the tensors are transformed again to proper dimension as required by EfficientNetB0 architecture. Following steps are performed:
\begin{itemize}
	\item The input tensor y is replicated along one dimension to create three copies, preparing it for subsequent operations.
	\item The expanded tensor is resized to a fixed size of (224, 224) using bilinear interpolation, ensuring consistent dimensions for neural network input.
	\item Any extra dimensions introduced during resizing are removed, returning the tensor to its original shape.
	\item The processed tensor is returned for further use, completing the preprocessing steps commonly applied to image data for deep learning tasks.
\end{itemize}
% \begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
% def expand_tensor_for_efficient_net(y) -> torch.Tensor:
%     input_tensor_expanded = y.expand(3, -1, -1)
%     resized_tensor = F.interpolate(input_tensor_expanded.unsqueeze(
%         0), size=(224, 224), mode='bilinear', align_corners=False)
